### 2016-08-27 reviews/reviewer1-response.md
Tweak r1 response

# Response to reviewer 1

We thank the reviewer for his or her detailed comments, which we have addressed
to improve the manuscript significantly.

>General remarks:
- I understand that the implementation of this work in OpenFOAM has demanded a lot of effort but could you please elaborate on what is new/novel in this work? An issue that has not been attacked/discovered/raised before or a numerical implementation which overcomes a known problem.

As noted in the "Novelty" file, this work presents some results that have either never been simulated or reported before:

1. Validating with Reynolds numbers more representative of a full-scale turbine, and two orders of magnitude higher than those used by Shamsoddin and Porte-Agel.
2. Reporting performance predictions, which has also not yet been done for a VAT in an ALM simulation.
3. Embedding the ALM inside of a RANS model for additional computational affordability.

The reviewer is correct that we had not sufficiently described this novelty in
the paper itself, and we have updated the introduction accordingly.

>- Minor comment: I'd rather use nondimentional units for variables such as time step

Figure 6 does use a time step that is technically nondimensional (time steps per rotor revolution).

>Pages 1-6: Again, very good introduction and gathering of introductory note,
but in my opinion, it should be shortened for a journal.

We agree and have shortened the introduction significantly.

TODO: Shorten introduction

>Page 7, Fig 3: It is taking a lot of room, perhaps you can make a smaller
schematic

We have made this figure smaller in the manuscript, a change that may be
retained by the typesetter.

>Page 9 fig 6: The Cp figure is an interesting one. But first of all, but is
there a particular reason why you have not shown Ct next to it (it's more common
this way than having Cd). Also, would it not be possible to have LES results on
top of both figures 6 and 11 for the cases you have simulated?

We use the overal rotor drag coefficient nomenclature ($C_D$) since this was
adopted by Paraschivoiu (2001). As noted, this is commonly called the thrust
coefficient for horizontal-axis turbines. We have added a note to the manuscript
to clarify this.

We have not plotted the LES results, since we only used LES for one tip speed
ratio, and for one turbine---it was mainly employed to assess wake predictions.

>Page 11, Line 42: I'm not quite sure if what you have stated about Smagorinsky
model is correct. It certainly is very easy to implement and gives nice results
in some cases, but we are all aware of its deficiencies. We set a fix Cs value
which means we add numerical dissipation to the whole computatiponal domain, but
this is not really accurate when dealing with cases with strong directional
homogeneity and anisotropy. Why didn't you choose a dynamic Smagorinsky or even
dynamic Lagrangian model? Besides, what is the Cs value you have used?

We are not trying to claim that the Smagorinsky model is the most accurate or
"the standard" model, but that is has been used in many other ALM papers, and
therefore a good baseline for comparison. The wording of the text has been
adjusted to reflect the notion that we are interested in this comparison to
baseline for turbine ALM modeling, and are by no means implying that this SGS
model is the most accurate or appropriate for this case. Certainly testing
additional SGS models will be part of future work, as is mentioned in the
conclusions.

The $C_S$ value used was OpenFOAM's default value of approximately 0.17. This
detail has been added to the text.

>Page 13, Fig 15: Following my last comment, figure to the right may very well
explain that the Smagorinsly model has been over dissipative thereby hampering
turbulent fluctuations which may have otherwise resulted in higher TKEs.

This is a valid point, and we have added a comment to the text accordingly. It
is important to note, however, that since we don't have near-wall boundary
layers, it may be that turbulence is only generated by larger-scale shear and
vortex breakdown, rather than by boundary layer separation or transition, which
would occur with a body-fitted mesh.

>Page 14, Line 27: Matching mean velocities is usually rather straightforward
(in non-presudospectral codes, sometimes even shutting down the whole SGS
modeling wouln'd make a difference as long as no solid walls are involved). But
higher order moments such as Reynolds stress tensor components are usually more
sensitive. Perhaps you can go more in depth with that part -- possibly compare
your results with other experiments or CFD simulations. Besides, please try and
overcome the relatively large discrepancies seen in figures 6 (high TSR), 9, 11.