### 2016-08-27 reviews/reviewer2-response.md
Make new figure to show computational domain

# Response to reviewer 2

>1. Pag1. Please insert the list of Symbols since there are a lot of
abbreviations in the text, a quick reference could be useful.

We have updated the manuscript with a nomenclature section.

>2. Pag. Pag3. Sect. 2. The parameter reduced frequency "k" is the key parameter
for unsteadiness considerations concerning VAT: some more detailed comments
should be given concerning the actual value of this parameter in the case study
numerically modelled. At the same time a comment should be given concerning the
ALM capabilities (and/or future improvements) in terms of unsteady aerodynamics.

>3. Pag4. Sect.4.  Is the force smoothing with the Gaussian function really
necessary? In P.-E. Réthoré and N. N. Sørensen, A discrete force allocation
algorithm, Wind Energy, 2012. Rethore and Sorensen showed how this spatial force
filter could be avoided with modified Rhie–Chow interpolation algorithm.

In our experience, especially with the finer LES grids, the force smoothing is
necessary. Without smoothing, the blade loading becomes quite "noisy" due to the
subsequent instability from sharper momentum source gradients. In this study we
did not investigate changes to the numerical algorithms used to solve the flow
field, but certainly future research may prove the Gaussian width can be
narrowed or eliminated if, for example, modified Rhie--Chow interpolation is
implemented in OpenFOAM.

>4. Pag6. Sect.7. Does the authors took into consideration that the ALM could be
in theory able to catch the Tip vortex effect without a correction model? i.e.
the vortex generated by the lifting of the blade line is already influencing the
angle of attack on the blade.

While tip vortices are produced by the ALM, and these do affect the local angle
of attack, we are not sure how easy it would be to create a ALM that resolves
tip effects totally accurately without any correction model. The issue may be
that since we do not enforce the no peneration BC through the blade, the tip
vortex can leave the blade too early, which causes incorrect angles of attack
close to the tip.

>5. Pag7. Sect.8-9 . Add a sketch of the domain boundaries with the position of
the turbine, and the flow direction. Figure 3 is not very effective in terms of
quantitative information on the actual implementation of the code.

This is a valid point. We have created a new figure to better describe the
computational domain and dimensions.

>6. Sect. 8-9 The details on the domain boundaries of the implemented case study
should be an opportunity for clarify if any blockage effect was present in the
reference experiments and if the same blockage conditions were adopted in the
numerical implementation.

The cross-section of the domain was set equal to match blockage, though the free
surface was not modeled. However, we did use a slip BC for the top as opposed to
a fixed tow speed condition, to get closer to the free surface condition. We
have added a small clarification regarding blockage matching to the text.

>7. Pag12. Sect.9.4 "The near-wake's mean velocity field at x/D=1 is shown in
Figure 12..." should be actually Figure 13

Thank you for catching this error. The manuscript has been corrected.