### 2018-03-29 reviews/reviewer2-response.md

# Response to reviewer 2

We thank the reviewer for his or her insightful comments, and feel that these
have helped improve the manuscript significantly.