### 2018-05-04 reviews/reviewer1-response.md
Start responses to reviewer 1

# Response to reviewer 1

We thank the reviewer for his or her detailed comments, which we have addressed
to improve the manuscript significantly.

>L. 111. Can you define lambda0?

This definition has been added to the theory section of the manuscript.

>Fig. 4 and 5. Can you include a line with the experimental value of Cp to appreciate how the model converges to it?

This is a reasonable suggestion. However, we feel that this may confuse the
readers with the difference between checking the independence of the model
parameters with the validation reported later.

>L. 228. Is the mesh uniform in the three directions? It can be quite relevant to state how the vertical resolution of the turbine is.

Yes, the mesh was uniform in all directions. This detail has been added to the
manuscript.

>L. 233. Define zero-subscript quantities.

Thanks for catching this. Is it now noted that $\Delta t_0$ is the time step
chosed to run the final simulations.

>L. 268. The drag and thrust can be considered differently to CFTs. Although for the whole rotor the drag can be considered as the thrust, i.e. as typically in a cylinder, it can bring to confusion when drag coefficient calculations on a blade are performed, as these are dependent on x- and y- forces, and not just on the former.

This is a valid point. Both are defined in different coordinate systems.
The clarify, we have attempted to separate blade element
and rotor drag using lowercase and uppercase subscripts, respectively.

>Fig. 6. The results showed in this figure are very promising. The overestimation at high tip speed ratios have been commonly found in literature as the authors pointed out. However, the reviewer questions if such effects can arise from the use of k-epsilon turbulence model. This model is usually not considered when wall effects are relevant in the analysed flow. Thus, the k-omega model is more used as in the author’s work from reference 10. Why did you not consider k-omega which would allow to a closer comparison to those blade-resolved results? At high tip speed ratios, the blades might undergo quite light dynamic stall, and maybe your DS model is over-predicting this value and triggering the Cd and Cp curves. Nevertheless, this is an almost unavoidable consequence of actuator line models. The reviewer suggest that the turbulence model may contribute more than expected by the authors in such Cp prediction.

We have doubts that the turbulence model is causing much difference, since
near actuator line elements, there is not a wall model. There could be an
effect from a difference in turbulence production in the model from shear
caused by the momentum source. We hypothesize that these discrepancies arise
from secondary effects, or potentially deficiencies in the input foil data,
both of which affect the simpler momentum and vortex models.

>Fig. 11. The results of power coefficient from the second case show that the model is quite case dependent. The results show that the model is not able to represent the physics about this turbine and thus the actual accuracy of the ALM cannot be verified without any additional figure that shows how the flow field is. Can you provide transversal plane contours showing the actual flow field generated by the ALM? Could this inaccuracy be again due to the k-epsilon model?

We have added a snapshot of the internal flow field and force vectors to give
and idea of the level of detail, or lack thereof, regarding local flow near
the blades. Unfortunately there is no mechanism for the ALM to generate flow
features like the pair of counter rotating vortices induced by dynamic stall.
This is a compromise inherent to the ALM, which we have clarified. It is
advantageous when attempting to simulate array-scale flows, but not so much
for individual device optimization, despite our validation efforts focusing
on the latter.

>Section 9.4. The presented LES results are quite disappointing due to the large differences against the RANS results which were very good. The reviewer has few concerns about this section:
>
Can you specify more details about the number of processors used? It is true that LES demands more computational resources than RANS but with the meshes used for the ALM it should be affordable to run LES.

We ran the LES simulations on 4 processors, a note that has been added to the
paper. It was still relatively affordable compared to the blade-resolved RANS.
We have noted a 2 order of magnitude drop in computational cost.

>Overall LES results are poor. The predicted Cp dropped to 20% which seems to be derived from the resolution of the flow field instead of the ALM itself. While a good agreement is found for the velocity wake as seen in Fig. 13 and 15(left), the turbulent kinetic energy results show an unexpected over damping of the velocity fluctuations. The use of LES is herein motivated to achieve a better resolution of the turbulent fluctuations by explicitly resolving the large-scale flow structures, and in the present manuscript the TKE has dropped dramatically. Fig. 14 and 15 (right) show a lack of flow unsteadiness attributed by the authors to the influence of the SGS model. The spatial discretisation is not described and it may be possible that depending on it, the scheme introduced an extra damping typically used in implicit LES wherein no SGS is used. Nevertheless, the reviewer would recommend to double-check the kinetic energy results as is quite confusing that the velocity wake prediction is greatly improved while the velocity fluctuations are almost gone.

One important note is that the SGS TKE is not tracked by the Smagorinsky
model, so there will be a lack of energy there.

It is not likely that the numerical schemes are dissipating the energy, since
all 2nd order schemes are used.

TODO: Check on TKE calculation from LES model and finish this.

>For further validation, the authors may consider to plot RANS vs LES at closer distance to the turbine, e.g. x/D=0.5 and study how the turbulent kinetic energy decay is in both models, or a transversal plane showing the generation of the vortices.